---
title: 'Argoid Analytics Assignment 2'
output: html_notebook
---
<b><h2>Reading the datafile

```{r}
transaction_data <- read.csv("C:/Users/sifat/Desktop/transaction_data.csv")
View(transaction_data)
data <- transaction_data

```

Exploratory data analysis</h2></b>

```{r}
glimpse(data)

```

<h4>Modifying the TranscationTime column for better analysis
```{r}
library(tidyr)
library(dplyr)
data <- data %>% separate(TransactionTime, c('week','month', 'date', 'hour','min','sec', 'zone','year'))
data$TransactionTime<- paste(data$month,data$date,data$year)
data <- data[,-c(3:10)]
head(data)

```

```{r}
summary(data)
```

The minimal value for NumberOfItemsPurchased and  CostPerItem  are negative . This could be because cancellation items transactions are captured in negative amount. So, we omitted the negative numbers.
Also, we have suspicious values for UserId and ItemCode.
So, we are ommiting negative values.
```{r}
data<-data[!(data$UserId == -1),]
data<-data[!(data$ItemCode == -1),]
data<-data[!(data$NumberOfItemsPurchased < 1),]
data<-data[!(data$CostPerItem < 1),]
summary(data)

```

Checking if there are NA values in any column
```{r}
sapply(data, function(x) any(is.na(x)))
```

```{r}
glimpse(data)
```

Due the elemination of unserirable rows the observations came down from 1,083,818 to 682,162. 

Checking the frequency tablbe for all the countries.

```{r}
table(data$Country)
```
The maximum orders came from United Kingdom.</h4>


<b><h2>HYPOTHESIS BUILDING</h2></b>

<h4>Restricting the data with maximum order that came from geographical unit the United Kingdom

```{r}
data = subset(data,Country == 'United Kingdom')
length(unique(data$TransactionId))
length(unique(data$UserId))
```

Converting the TransactionTime column in date format using anytime pacakage.
Also, converting some other columns into factors for better analysis.
Creating a column Total price by multiplying CostPerItem and NumberOfItemsPurchased
```{r}
library(anytime)
data$Country = as.factor(data$Country)
data$TransactionId = as.factor(data$TransactionId)
data$ItemCode = as.factor(data$ItemCode)
data$UserId = as.factor(data$UserId)
data$TransactionTime = as.factor(data$TransactionTime)

q <- anydate(data$TransactionTime)
data$TransactionTime <- q
total <- df$CostPerItem *df$NumberOfItemsPurchased
data$Totalprice <- total
```

```{r}
glimpse(data)
```

Analysing the dataset using RFM Technique

```{r}
library(dplyr)
rmf = data %>% 
  group_by(UserId) %>%
  summarise(last_purchase = max(TransactionTime),
            recency = as.double(difftime('2019-01-01',
                                            as.Date(last_purchase,origin= 
                                                      '2018-02-12 '),
                                              units = 'day')),
            frequency = n_distinct(TransactionId),
            momentary = sum(Totalprice)
            )
head(rmf, n=20)

rmf<-rmf[!(rmf$recency < 0),]
head(rmf, n= 20)

```

```{r}
glimpse(rmf)
```

```{r}
rmf2 = rmf %>%
      mutate(one_time = ifelse(frequency == 1 , "One-timer", "More than once"),
             avg_per_purchase = momentary/frequency) %>%
      group_by(one_time)
head(rmf2, n= 30)

```
```{r}
library(ggplot2)
library(plotly)

pl <- ggplot(rmf2 , aes(x= one_time, y=momentary))

pl2 <- pl+geom_bar(stat = 'identity',fill = '#f54287', alpha= 0.95)

pl3 <- pl2 + xlab('How many times')+ylab('Momentary')

gpl<- ggplotly(pl3)

print( gpl )
```
Creation of R, F, M rank

```{r}
rmf2$r = cut(rmf2$recency,5, F)
rmf2$f = cut (rmf2$frequency, 5, F)
rmf2$m = cut(rmf2$momentary, 5 , F)

```

Review top 20
```{r}
rmf3 = rmf2[with(rmf2, order(-r, -f, -m)),]
head(rmf3, n=20)
```

```{r}
groupRFM = count(rmf3, r,f,m)
groupRFM = rmf3$r*100 +rmf3$f*10+rmf3$m
rmf3 = data.frame(rmf3, groupRFM)
head(rmf3)

```
Giving scored to the customers according to their RFM values

```{r}
rmf3$score <- rmf3$r+rmf3$f+rmf3$m
countRFM = count(rmf3, score)
countRFM
```
RFM scores based on the Quartile (Distribustion based)


```{r}
summary(rmf3$recency)
```

```{r}
rmf3$segment_recency = 'NA'
rmf3$segment_recency[which(rmf$recency > 191.7708)]= 'Inactive'
rmf3$segment_recency[which(rmf3$recency > 92.7708 & rmf3$recency<= 191.7708)] <- 'Temporary idle'
rmf3$segment_recency[which(rmf3$recency > 26.7708 & rmf3$recency<= 92.7708)] <- 'Frequent'
rmf3$segment_recency[which(rmf3$recency<= 26.7708)] <- 'Highly Active'
rfm3_recency = rmf3
head(rfm3_recency)

```

```{r}
customerSegmentationR = count(rmf3, segment_recency)
customerSegmentationR
```

```{r}
pl <- ggplot(customerSegmentationR, aes(x=segment_recency, y = n))

pl2 <- pl+geom_bar(stat = 'identity',fill = '#fcba03', alpha= 0.95)

pl3 <- pl2 + xlab('Customer status')+ylab('count')

gpl<- ggplotly(pl3)

print( gpl )

```

```{r}
summary(rmf3$frequency)
```

```{r}
rmf3$segment_frequency = 'NA'
rmf3$segment_frequency[which(rmf$frequency <1)]= 'Potential'
rmf3$segment_frequency[which(rmf3$frequency >=1 )] <- 'C. Returing customer'
rmf3$segment_frequency[which(rmf3$frequency >= 2 )] <- 'B. Regular'
rmf3$segment_frequency[which(rmf3$frequency >= 5)] <- 'A. Stellar customer'
rfm3_frequency = rmf3
head(rfm3_frequency, n= 20)
```

```{r}
customerSegmentationF = count(rmf3, segment_frequency)
customerSegmentationF

```

```{r}
pl <- ggplot(customerSegmentationF, aes(x=segment_frequency, y = n))

pl2 <- pl+geom_bar(stat = 'identity',fill = '#34eb43', alpha= 0.95)

pl3 <- pl2 + xlab('Customer status')+ylab('count')

gpl<- ggplotly(pl3)

print( gpl )
```

```{r}
summary(rmf3$momentary)
```

```{r}
rmf3$segment_momentary = 'NA'
rmf3$segment_momentary[which(rmf$momentary < 788)]= 'worst'
rmf3$segment_momentary[which(rmf3$momentary >= 788 & rmf3$momentary<  1622 )] <- 'Top 75%'
rmf3$segment_momentary[which(rmf3$momentary >=  1622 & rmf3$momentary< 3360)] <- 'Top 50%'
rmf3$segment_momentary[which(rmf3$momentary>= 3360)] <- 'Best'
rfm3_momentary = rmf3
head(rfm3_momentary, n = 20)
```
```{r}
customerSegmentationM= count(rmf3, segment_momentary)
customerSegmentationM
```

```{r}
pl <- ggplot(customerSegmentationF, aes(x=segment_frequency, y = n))

pl2 <- pl+geom_bar(stat = 'identity',fill = '#ebc634', alpha= 0.95)

pl3 <- pl2 + xlab('Customer status')+ylab('count')

gpl<- ggplotly(pl3)

print( gpl )
```
</h4>
```{r}
write.csv(rmf3, "RFMresult.csv")

```

